---
title: "Homework 2"
author: "Susan Davis"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:**  1-3 (Pgs. 49-50); 2-4 (Pg. 51); 2,4 (Pg. 54); 1-4 (Pg. 58)

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a HTML file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining. 

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
library(nycflights13)
```

# Chapter 3 (Pgs. 49-50)

##  Exercise 1
1. Find all flights that:
a. Had an arrival delay of two or more hours
b. Flew to Houston (IAH or HOU)
c. Were operated by United, American, or Delta
d. Departed in summer (July, August, and September)
e. Arrived more than two hours late, but didn’t leave late
f. Were delayed by at least an hour, but made up over 30
minutes in flight
g. Departed between midnight and 6 a.m. (inclusive)

```{r}
head(flights)
?flights
filter(flights, dep_delay >= 2)
```
```{r}
filter(flights, dest == 'IAH' | dest == 'HOU')
```
```{r}
filter(flights, carrier == 'AA'| carrier ==  'UA'| carrier ==  'DL' )
```
```{r}
filter(flights, month%in% c(7:9))
```

##  Exercise 2
```{r}
#
```

##  Exercise 3
```{r}
#
```

# Chapter 3 (Pg. 51)

##  Exercise 2
```{r}
#
```

##  Exercise 3
```{r}
#
```

##  Exercise 4
```{r}
#
```

# Chapter 3 (Pg. 54)

##  Exercise 2
```{r}
#
```

##  Exercise 4
```{r}
#
```

# Chapter 3 (Pg. 58)

##  Exercise 1
```{r}
#
```

##  Exercise 2
```{r}
#
```

##  Exercise 3
```{r}
#
```

##  Exercise 4
```{r}
#
```